### 2021-03-08 FrontBackmatter/Declaration.tex
Lots of small improvements

%*******************************************************
% Declaration
%*******************************************************
\pdfbookmark[0]{Declaration}{declaration}
\chapter*{Declaration}
\thispagestyle{empty}
Put your declaration here.
\bigskip

>>>>

%*******************************************************
% Declaration
%*******************************************************
\pdfbookmark[0]{Declaration}{declaration}
\chapter*{Declaration}
\thispagestyle{empty}
I hereby declare that this thesis has not been and will not be, submitted in whole or in part to another University for the award of any other degree.
Except where indicated by specific stated in the text, this thesis was composed by myself and the work contained therein in my own.
\bigskip

### 2021-03-08 ch/2-related-work.tex
Lots of small improvements

\citet{arjovsky2019invariant} tackle a similar problem, but take a very different approach,
which they term ``Invariant Risk Minimization'', in contrast to the default method Empirical Risk Minimization.
As with \citet{kim2019learning},
the goal is to ignore spurious correlations that only appear in the very imperfect training set,
but not in the test set;
and they also perform experiments on Coloured MNIST (see above).
They formalise the problem as one of different environments $e\in E$,
where each environment is a different, biased view of the same underlying data distribution.
The idea is to train a predictor that is simultaneously optimal for all the environments,
with the expectation that this generalises to the test set.
This idea is formulated as a (intractable) bi-level, constrained optimization problem,
which they approximate with a tractable regularised optimization.

\citet{creager2020environment} build directly on \citet{arjovsky2019invariant}.
They aim to address the limitation that the environments -- that one wants to be invariant to -- have to be pre-defined.
Their contribution is to infer these environments instead,
based on identifying which environment splits would most negatively affect an ERM classifier.
Their experiments show that this can even approve upon hand-crafted environments;
experiments are performed on a synthetic dataset and Coloured MNIST.
In contrast to \citet{kim2019learning} and \citet{arjovsky2019invariant},
the authors explicitly point out the connection to traditional fairness methods.

>>>>

\citet{arjovsky2019invariant} tackle a similar problem, but take a very different approach,
which they term ``Invariant Risk Minimisation'', in contrast to the default method \acf{ERM}.
As with \citet{kim2019learning},
the goal is to ignore spurious correlations that only appear in the very imperfect training set,
but not in the test set;
and they also perform experiments on Coloured MNIST (see above).
They formalise the problem as one of different environments $e\in E$,
where each environment is a different, biased view of the same underlying data distribution.
The idea is to train a predictor that is simultaneously optimal for all the environments,
with the expectation that this generalises to the test set.
This idea is formulated as a (intractable) bi-level, constrained optimization problem,
which they approximate with a tractable regularised optimization.

\citet{creager2020environment} build directly on \citet{arjovsky2019invariant}.
They aim to address the limitation that the environments -- that one wants to be invariant to -- have to be pre-defined.
Their contribution is to infer these environments instead,
based on identifying which environment splits would most negatively affect an \ac{ERM} classifier.
Their experiments show that this can even approve upon hand-crafted environments;
experiments are performed on a synthetic dataset and Coloured MNIST.
In contrast to \citet{kim2019learning} and \citet{arjovsky2019invariant},
the authors explicitly point out the connection to traditional fairness methods.

### 2021-03-08 ch/4-structure.tex
Lots of small improvements

\section{Using target labels to mitigate label bias}\label{sec:target-labels}
\Chapref{ch:paper1} is predominantly concerned with label bias.
The main idea is that we make use of pseudo labels (or \emph{target labels})
to implicitly learn from a \emph{balanced} dataset,
in which \(y\perp s\) holds, and which thus satisfies \acf{DP}.
This falls under the area of \emph{fair classifiers} discussed in \secref{fair-classifiers}

We can interpret the contributions of this chapter in two ways.
The first corresponds to the definition-centric view of dataset bias,
and the second to the ground-truth-centric view.
\begin{enumerate}
  \item
    We can say that the classifier should satisfy demographic parity in its predictions,
    and learning from a balanced training set is just one particular way to achieve this.
    In this view, the pseudo labels have no deeper meaning and are just a computational trick.
  \item
    We can see the training set as a corrupted version of a true dataset, which is balanced (\(y\perp s\)),
    and so, by learning from these pseudo labels, we are simply approximating the true dataset.
    However, we do not actually have access to the true dataset; we only know that it is balanced.
    In order to evaluate the trained model, we compute fairness metrics with respect to \ac{DP}.
\end{enumerate}
Within the chapter, we sometimes jump between these two views.

@@

\section{Using a representative set to overcome severe sampling bias}\label{sec:nifr}
In the setting from the previous chapter,
labels were untrustworthy because they had been flipped;
a phenomenon we referred to as \emph{label bias}.
However, flipped labels is not the only way that labels can become untrustworthy.
Another way is \emph{sampling bias}, which is the subject of \twochaprefs{ch:paper2}{ch:paper3}.

@@

The method, proposed in the previous chapter,
is not able to deal with such a dataset bias as we can easily see:
Say, \emph{smiling} corresponds to \(y=1\) and \emph{not smiling} to \(y=0\);
furthermore, let red hair correspond to \(s=1\), black hair to \(s=0\), and all other hair colours to \(s=2\).
Then, the problem with the described dataset is, that it mostly consists of samples with \(y=0\wedge s=0\)
and those with \(y=1\wedge s=1\).
If we call \(P(y=1|s=s')\) the acceptance rate,
then the problem can be described as one of very different acceptance rates in the hair colour groups given by \(s\).
This is the problem tackled in the previous chapter,
and yet, if we were to equalize the acceptance rates with the method there,
the result would be very incorrect.
The issue is that we would treat the labels as incorrect, when in truth, they are correct.

@@

A parallel to the previous chapter is that the method makes use of side information
(in this case the representative set)
in order to overcome the bias in the training set.

The method presented in the chapter is based on the idea of \emph{null-sampling},
which refers to zeroing out part of an encoding,
and then reconstructing the modified encoding as if it were a normal encoding.
Before we apply null-sampling, an encoding of the input \(x\) is learned that is split into two parts:
\(z_u\), which has no information about the spurious attribute \(s\),
and \(z_b\), which has all information needed to reconstruct \(x\) that is not contained in \(z_u\).
\(z_u\) is ensured to be not predictive of \(s\) via adversarial training.
During null-sampling, \(z_b\) is zeroed out, and after decoding it,
we obtain an invariant representation \emph{in the data domain}.
The fact that it is in the data domain makes it interpretable.

@@

\section{Using an unlabelled deployment set to overcome sam\-p\-ling bias}\label{sec:zsf}
A shortcoming of the approach from the previous chapter
is its reliance on a representative set which has \(s\) labels.
As discussed, it is necessary to make use of \emph{some} kind of side information,
but perhaps we can relax some requirements.
In particular, while it is already easier to collect data without \(y\) labels (but with \(s\) labels),
it is \emph{even} easier to collect data without \emph{any} label.
Thus, requiring only an unlabelled context set would improve the applicability of the method.
This chapter presents an approach based on that idea.

The setting is very similar to the previous chapter:
the training set suffers from severe sampling bias, but we have access to a (mostly) unbiased \emph{deployment set}.
The idea is that the deployment set corresponds to the setting in which the model is meant to be deployed.
The change is that this additional set may be completely unlabelled,
but in exchange, we have some stronger requirements for the training set.
Namely, the holes left by the sampling bias may not be so numerous
as to make \(s\) and \(y\) completely indistinguishable.
In the previous chapter, the example of Coloured MNIST had a training set
where there was a strict one-to-one mapping of colour and digit;
but this kind of blending of \(s\) and \(y\) into one is not the focus of this chapter.
The focus is instead on a setting where the training set lacks certain combinations of \(s\) and \(y\),
which results in poor predictions for these combinations on the test set (or deployment setting)
where those combinations do occur.
We refer to these missing combinations as \emph{invisible demographics}.
The value of \(s\) acts as a spurious attribute and should be ignored in order to make accurate predictions.

As in the previous chapter, the first step is to train a neural network to produce an invariant representation.
The second step is then to train a classifier on that.
The invariant representation is trained by doing distribution matching between the training set and the deployment set.
Finally, the training set is transformed into the invariant representation,
to serve as training set for the classifier.

>>>>

\section{Using target labels to mitigate label bias}\label{sec:target-labels}
\citet{kehrenberg2020tuning} (\chapref{ch:paper1}) is predominantly concerned with label bias.
The main idea is that we make use of pseudo labels (or \emph{target labels})
to implicitly learn from a \emph{balanced} dataset,
in which \(y\perp s\) holds, and which thus satisfies \acf{DP}.
This falls under the area of \emph{fair classifiers} discussed in \secref{fair-classifiers}

We can interpret the contributions of this publication in two ways.
The first corresponds to the definition-centric view of dataset bias,
and the second to the ground-truth-centric view.
\begin{enumerate}
  \item
    We can say that the classifier should satisfy demographic parity in its predictions,
    and learning from a balanced training set is just one particular way to achieve this.
    In this view, the pseudo labels have no deeper meaning and are just a computational trick.
  \item
    We can see the training set as a corrupted version of a true dataset, which is balanced (\(y\perp s\)),
    and so, by learning from these pseudo labels, we are simply approximating the true dataset.
    However, we do not actually have access to the true dataset; we only know that it is balanced.
    In order to evaluate the trained model, we compute fairness metrics with respect to \ac{DP}.
\end{enumerate}
Within the paper, we sometimes jump between these two views.

@@

\section{Using a representative set to overcome severe sampling bias}\label{sec:nifr}
In the setting from the previous paper \citep{kehrenberg2020tuning},
labels were untrustworthy because they had been flipped;
a phenomenon we referred to as \emph{label bias}.
However, flipped labels is not the only way that labels can become untrustworthy.
Another way is \emph{sampling bias}, which is the subject of \citet{kehrenberg2020nullsampling} and \citet{kehrenberg2020zeroshot} (\twochaprefs{ch:paper2}{ch:paper3}).

@@

The method, proposed in the previous paper (\chapref{ch:paper1}),
is not able to deal with such a dataset bias as we can easily see:
Say, \emph{smiling} corresponds to \(y=1\) and \emph{not smiling} to \(y=0\);
furthermore, let red hair correspond to \(s=1\), black hair to \(s=0\), and all other hair colours to \(s=2\).
Then, the problem with the described dataset is, that it mostly consists of samples with \(y=0\wedge s=0\)
and those with \(y=1\wedge s=1\).
If we call \(P(y=1|s=s')\) the acceptance rate,
then the problem can be described as one of very different acceptance rates in the hair colour groups given by \(s\).
This is the problem tackled in the previous paper,
and yet, if we were to equalize the acceptance rates with the method there,
the result would be very incorrect.
The issue is that we would treat the labels as incorrect, when in truth, they are correct.

@@

A parallel to the previous paper is that the method makes use of side information
(in this case the representative set)
in order to overcome the bias in the training set.

The method presented in \citet{kehrenberg2020nullsampling} (\chapref{ch:paper2}) is based on the idea of \emph{null-sampling},
which refers to zeroing out part of an encoding,
and then reconstructing the modified encoding as if it were a normal encoding.
Before we apply null-sampling, an encoding of the input \(x\) is learned that is split into two parts:
\(z_u\), which has no information about the spurious attribute \(s\),
and \(z_b\), which has all information needed to reconstruct \(x\) that is not contained in \(z_u\).
\(z_u\) is ensured to be not predictive of \(s\) via adversarial training.
During null-sampling, \(z_b\) is zeroed out, and after decoding it,
we obtain an invariant representation \emph{in the data domain}.
The fact that it is in the data domain makes it interpretable.

@@

\section{Using an unlabelled deployment set to overcome sam\-p\-ling bias}\label{sec:zsf}
A shortcoming of the approach from the previous publication \citep{kehrenberg2020nullsampling}
is its reliance on a representative set which has \(s\) labels.
As discussed, it is necessary to make use of \emph{some} kind of side information,
but perhaps we can relax some requirements.
In particular, while it is already easier to collect data without \(y\) labels (but with \(s\) labels),
it is \emph{even} easier to collect data without \emph{any} label.
Thus, requiring only an unlabelled context set would improve the applicability of the method.
\citet{kehrenberg2020zeroshot} (\chapref{ch:paper3}) presents an approach based on that idea.

The setting is very similar to the previous publication \citep{kehrenberg2020nullsampling}:
the training set suffers from severe sampling bias, but we have access to a (mostly) unbiased \emph{deployment set}.
The idea is that the deployment set corresponds to the setting in which the model is meant to be deployed.
The change is that this additional set may be completely unlabelled,
but in exchange, we have some stronger requirements for the training set.
Namely, the holes left by the sampling bias may not be so numerous
as to make \(s\) and \(y\) completely indistinguishable.
In the previous paper, the example of Coloured MNIST had a training set
where there was a strict one-to-one mapping of colour and digit;
but this kind of blending of \(s\) and \(y\) into one is not the focus of this paper.
The focus is instead on a setting where the training set lacks certain combinations of \(s\) and \(y\),
which results in poor predictions for these combinations on the test set (or deployment setting)
where those combinations do occur.
We refer to these missing combinations as \emph{invisible demographics}.
The value of \(s\) acts as a spurious attribute and should be ignored in order to make accurate predictions.

As in the previous paper, the first step is to train a neural network to produce an invariant representation.
The second step is then to train a classifier on that.
The invariant representation is trained by doing distribution matching between the training set and the deployment set.
Finally, the training set is transformed into the invariant representation,
to serve as training set for the classifier.

### 2021-03-08 paper2/2-related-work.tex
Lots of small improvements

\subsection{Learning fair representations.}
%As we have alluded to, the goal of producing invariant representations is similar to that of producing \emph{fair} representations.
% In fairness problems, there is usually a \emph{sensitive attribute}, $s$ (for example, gender or race),
% that should not be used to make decisions.
Given a sensitive attribute $s$ (for example, gender or race) and inputs $\bm{x}$,
a fair representation $\bm{z}$ of $\bm{x}$ is then one for which $\bm{z} \perp s$ holds,
while ideally also being predictive of the class label $y$.
\citet{zemel2013learning} was the first to propose the learning of fair representations which allow for transfer to new classification tasks.
More recent methods are often based on variational autoencoders (VAEs)~\citep{kingma2013auto,louizos2016variational,edwards2016censoring,beutel2017data}.
The achieved fairness of the representation can be measured with various fairness metrics.
These measure, however, usually how fair the predictions of a classifier are
and not how fair a representation is.
% defined with respect to predictions and not representations.

% There is not one metric universally best-suited to measuring the invariance (or fairness) of representations.
The appropriate measure of fairness for a given task is domain-specific \citep{liu2018delayed}
and there is often not a universally accepted measure.
However, \emph{Demographic Parity} is the most widely used~\citep{louizos2016variational,edwards2016censoring,beutel2017data}.
Demographic Parity demands $\hat{y} \perp s$ where $\hat{y}$ refers to the predictions of the classifier.
In the context of fair representations, we measure the Demographic Parity of a downstream classifier, $f(\cdot )$, which is trained on the representation $z$ i.e.  $f: Z \to \hat{Y}$.

>>>>

\subsection{Learning fair representations.}
%As we have alluded to, the goal of producing invariant representations is similar to that of producing \emph{fair} representations.
% In fairness problems, there is usually a \emph{sensitive attribute}, $s$ (for example, gender or race),
% that should not be used to make decisions.
Given a sensitive attribute $s$ (for example, gender or race) and inputs $\bm{x}$,
a fair representation $\bm{z}$ of $\bm{x}$ is then one for which $\bm{z} \perp s$ holds,
while ideally also being predictive of the class label $y$.
\citet{zemel2013learning} was the first to propose the learning of fair representations which allow for transfer to new classification tasks.
More recent methods are often based on \acfp{VAE}~\citep{kingma2013auto,louizos2016variational,edwards2016censoring,beutel2017data}.
The achieved fairness of the representation can be measured with various fairness metrics.
These measure, however, usually how fair the predictions of a classifier are
and not how fair a representation is.
% defined with respect to predictions and not representations.

% There is not one metric universally best-suited to measuring the invariance (or fairness) of representations.
The appropriate measure of fairness for a given task is domain-specific \citep{liu2018delayed}
and there is often not a universally accepted measure.
However, \emph{Demographic Parity} is the most widely used~\citep{louizos2016variational,edwards2016censoring,beutel2017data}.
Demographic Parity demands $\hat{y} \perp s$ where $\hat{y}$ refers to the predictions of the classifier.
In the context of fair representations, we measure the Demographic Parity of a downstream classifier, $f(\cdot )$, which is trained on the representation $z$, i.e.\  $f: Z \to \hat{Y}$.
